---
title: "HW11_JLingle_MSDS6306"
author: "Jason Lingle"
date: "April 10, 2018"
output: html_document
---

## Doing Data Science HW 11: Time Series

###1.Warm Up: Brief Financial Data (20%)
a. Natively in R, you have access to sample data sets of prominent stocks over time.
We'll be using EuStockMarkets for this question. Type help(EuStockMarkets) to learn
more. From these data, pull specifically the DAX index. For all questions in this
assignment, you're welcome to normalize (or don't!) how you see fit, but, if you
choose to, please document what you're doing and why for the grader. It's not
necessary for the purpose of this assignment.

```{r, echo=FALSE}
help(EuStockMarkets)
```

b. These are annual European Stock Data from 1990 onward. Create a rudimentary plot
of the data. Make the line blue. Give an informative title. Label the axes accurately.
In 1997, an event happened you want to indicate; add a vertical red line to your plot
which divides pre-1997 and post-1997 information.

```{r, echo=TRUE}
plot(EuStockMarkets[,"DAX"], col="blue", xlab="Year", ylab="DAX Stock Price", main="German DAX Stock Prices Per Year")
abline(v="1997", col="red")
```

c. Decompose the time series into its components (i.e., trend, seasonality, random). Keep
in mind that this is a multiplicative model you want. Create a plot of all decomposed
components. As before, make all lines blue and have a vertical divider at the year
1997.

```{r, echo=TRUE}
decomposedEU <- decompose(EuStockMarkets[,"DAX"], type="mult")
plot(decomposedEU, col="blue")
abline(v="1997", col="red")
```

###2.Temperature Data (40%)
a. Using the maxtemp dataset granted by loading fpp2, there are maximum annual
temperature data in Celsius. For more information, use help(maxtemp). To see what
you're looking at, execute the command in 'Examples' in the help document.

```{r, echo=TRUE}
library(fpp2)
help(maxtemp)
autoplot(maxtemp)
```

b. We are only concerned with information after 1990. Please eliminate unwanted
information or subset information we care about.

```{r, echo=TRUE}
maxtempts <- window(maxtemp,start=1990)
```

c. Utilize SES to predict the next five years of maximum temperatures in Melbourne.
Plot this information, including the prior information and the forecast. Add the
predicted value line across 1990-present as a separate line, preferably blue. So, to
review, you should have your fit, the predicted value line overlaying it, and a forecast
through 2021, all on one axis. Find the AICc of this fitted model. You will use that
information later.

```{r, echo=TRUE}
fit1 <- ses(maxtempts, h=5)
plot(maxtempts, ylab="Maximum Temperatures", main="Maximum Temperatures in Melbourne Since 1990")
plot(fit1, col="blue")
```

d. Now use a damped Holt's linear trend to also predict out five years. Make sure
initial="optimal." As above, create a similar plot to 1C, but use the Holt fit instead.

```{r, echo=TRUE}
fit2 <- holt(maxtempts, initial="optimal", h=5)
fit2$model
plot(maxtempts, ylab="Maximum Temperatures", main="Maximum Temperatures in Melbourne Since 1990")
plot(fit2, col="blue")
```

e. Compare the AICc of the ses() and holt() models. Which model is better here?

```{r, echo=TRUE}
accuracy(fit1)
accuracy(fit2)
```


###3.The Wands Choose the Wizard (40%)
a. Utilize the dygraphs library. Read in both Unit11TimeSeries_Ollivander and
_Gregorovitch.csv as two different data frames. They do not have headers, so make
sure you account for that. This is a time series of Wands sold over years.

```{r, echo=TRUE}
library(dygraphs)
setwd("C:/MSDS6306/")
Ollivander <- read.csv("Unit11TimeSeries_Ollivander.csv", sep = ",", header=FALSE)
colnames(Ollivander) <- c("Year","Wands_Sold")
Gregor  <-  read.csv("Unit11TimeSeries_Gregorovitch.csv", sep = ",", header=FALSE)
colnames(Gregor) <- c("Year","Wands_Sold")
```


b. You don't have your information in the proper format! In both data sets, you'll need to
first convert the date-like variable to an actual Date class.

```{r, echo=TRUE}
Gregor$Year <- as.Date(Gregor$Year, format = "%m/%d/%Y")
Ollivander$Year <- as.Date(Ollivander$Year, format = "%m/%d/%Y")
```


c. Use the library xts (and the xts() function in it) to make each data frame an xts object
(effectively, a time series). You'll want to order.by the Date variable.

```{r, echo=TRUE}
library(xts)
Gregorxts <- xts(x=Gregor, order.by=Gregor$Year)
Ollivanderxts <- xts(x=Ollivander, order.by=Ollivander$Year)
```

d. Bind the two xts objects together and create a dygraph from it. Utilize the help() index
if you're stuck.
. Give an effective title and x/y axes.
. Label each Series (via dySeries) to be the appropriate wand-maker. So, one
line should create a label for Ollivander and the other for Gregorovitch.
. Stack this graph and modify the two lines to be different colors (and not the
default ones!) Any colors are fine, but make sure they're visible and that
Ollivander is a different color than Gregorovitch.
. Activate a range selector and make it big enough to view.
. Use dyShading to illuminate approximately when Voldemort was revived
and at-large: between 1995 to 1999.
. Enable Highlighting on the graph, so mousing over a line bolds it.

```{r, echo=TRUE}
Wands <- cbind(Gregorxts$Wands_Sold, Ollivanderxts$Wands_Sold)
colnames(Wands) <- c("Gregorovitch","Ollivander")
dygraph(Wands, xlab="Year",ylab="Wands Sold", main="Wands Sold Per Year")  %>% dyRangeSelector() %>% dyOptions(colors = RColorBrewer::brewer.pal(3, "Set1")) %>% dyShading(from = "1995-1-1", to = "1999-1-1")
```